{\rtf1\mac\ansicpg10000\cocoartf824\cocoasubrtf410
{\fonttbl\f0\fswiss\fcharset77 Helvetica;}
{\colortbl;\red255\green255\blue255;}
\margl1440\margr1440\vieww10500\viewh9600\viewkind0
\deftab720
\pard\pardeftab720\ri720\ql\qnatural

\f0\fs26 \cf0 1. Short introduction.\
\
2. Optical requirements. Youri\
\
3. Physical environment. Youri.\
	-Location and spatial constraints\
	-Physical environment, manetic fields, etc.\
	-Limits on materials\
	\
4. Overall design. Youri.\
\
5. Simulation results. Alex.\
\pard\tx560\tx1120\tx1680\tx2240\tx2800\tx3360\tx3920\tx4480\tx5040\tx5600\tx6160\tx6720\ql\qnatural\pardirnatural
\cf0     	-Choice of gas\
	-Nphe in simplest case - to show that it is enough\
	-Expected PMT window effect on Nphe numbers\
	-Possible distortion sources, - Magnetic Field, target lenght,\
	-Target position, mirror surface and position inperfection.\
	-Choice of WC window size.\
	-Shape of PMT window and shape of WC.\
	-Estimate of Nphe and efficiency in realistic case.\
\
\pard\pardeftab720\ri720\ql\qnatural
\cf0 \
6. Details of optical design and construction. Youri\
\
7. Mirror prototyping activity and plans. Youri\
\
7. Details of light collection cones design and construction. \
\
8. Details of PMT considerations. Valeri\
	Tubes and responses.\
	Ambient magnetic fields.\
	Details of  shielding design and characteristics.\
	PMT prototyping activity.\
\
9. Details of mechanical design. Dave \
	Current activities and near term plans.\
\
10. Division of responsibilities. TBD\
\
11. Planned activities. TBD\
\
12. Cost table.  Youri and Paul\
}