%\section{Bibliography}
%\addcontentsline{toc}{section}{Bibliography}

\begin{thebibliography}{99}

\bibitem{upgrade} Jefferson Lab 12 GeV Upgrade Project, http://www.jlab.org/12GeV/.

\bibitem{r2} E.S.Smith et al., ``The time-of-flight system for CLAS'',  Nucl. Instr. and Meth. A \textbf{432},265-298~(1999).


\bibitem{Baturin:2005} V.Baturin {\it et al.}, ``Time-of-flight resolution of scintillating counters with
Burle 85001 micro-channel plate photo-multipliers  in comparison with Hamamatsu R2083 '',
Nucl. Instr. and Meth. A \textbf{562},327-337~(2006).


\bibitem{r1} V.N.Batourine, W.Kim, D.M.Nekrasov, K.Park, B.Shin, E.S.Smith
 and S.S.Stepanyan,
''Measurements of PMT time resolution at Kyungpook National University'',
CLAS-NOTE~2004-16(JLAB,2004),

\bibitem{cn200439} V.N.Batourine, W.Kim, S.Majewsky, V.Popov, E.S.Smith, 
D.Son and C.Zorn ,
''Studies of time resolution of the Burle 85001 micro-channel 
plate photo-multipliers in comparison with standard PMTs'', CLAS-NOTE~2004-39(JLAB,2004).

\bibitem{llg}  V.Batourine, V.Burkert, W.Kim, G.S.Mutchler,  D.Nekrasov and E.S. Smith
''Status Report on the studies at Kyungpook National University in 2005'', 
CLAS-NOTE~2005-3(JLAB,15-Jul-2005).

\bibitem{cn85011}  F.Barbosa, V.Batourine, V.Burkert, C.Cuevas,
 L.Elouadrhiri, W.Kim, S.Majewski, G.S.Mutchler, V.Popov, E.S.Smith and C.Zorn,
''Status and further steps towards the CLAS12 start-counter'',CLAS-NOTE~2006-011(JLAB,01-Aug-2006)





\bibitem{mutch} G.Mutchler and Y.Sharabian, ``Wrapping Tests and Monte Carlo 
evaluation of a new Highly Segmented CLAS Start Counter''. CLAS-NOTE~2005-004.




\bibitem{pa85011}  F.Barbosa and V.Popov
 `` The assembly of MCP PM Burle85011 with the on-board preamplifier'', 
to be published as CLAS NOTE.
\verb|http://www.jlab.org/~baturin/Burle85011WideDynamicRangePMTAssembly.pdf|
\verb|http://www.jlab.org/~baturin/Burle85011_HV.pdf|
\verb|http://www.jlab.org/~baturin/Burle85011_schematics.pdf|


\bibitem{kich} H.Kichimi, H.Sagawa, Y.Yoshimura, T.Kishida, N.Tamura,
 H.Suzuki and T.Taguchi,
 ``Timing characteristics of micro-channel plate and fine mesh photomultiplier
 tubes in a 1-Tesla field'', Nucl.\ Instrum.\ Meth.\ A {\bf 325},451(1993).

\bibitem{vavra} J.Va'vra, ``Photon detectors'', SLAC,Japan 2005,\\ 
\verb|rd.kek.jp/slides/20051205/photon_detectors.pdf|



\bibitem{fmtr01} M.Bonesini et al., ``Behaviour in high magnetic fields of fine-mesh photodetectors for fast time-of-flight detectors'',
  Nucl. Instr. and Meth. A \textbf{567},200-204~(2006).




\bibitem{foucher} M.Foucher, M.Herndon, H.Jawahery, 
``A study of the timing characteristics of scintillator counters with fine-mesh photomultiplier tubes'',
 Nucl. Instr. and Meth. A \textbf{374},57-62~(1996).


\bibitem{csn1} G.Cecchet,  ``Timing counter status'',CSNI, Assisi 2004,\\
\verb|www.infn.it/csn1/riunioni/agenda/20-09-2004/cecchet_timing_counter.pdf|

%\bibitem{atlas} \\
%\verb|http://atlas.ifae.es/~cortina/biblio/Experiments/SiPM_slac.pdf|




\bibitem{dolgo} B.Dolgoshein, ``Large area Silicon Photomultiplier: Performance and Applications'',
4th ICNDP, Beaune-2005,\\
 \verb|www.jlab.org/~rondon/sane/mtg8/Dolgoshein.pdf|

\bibitem{apd} Alfred Q.R. Baron.    ``Avalanche Diodes (APD) Detectors: Introduction and Recent Advances'',
Harima RIKEN \& Spring-8/JASRI, International Symposium on the Development of Detectors
for Particle, Astro-Particle, and Synchrotron Radiation Experiments, SLAC, Stanford, April 2006,\\
\verb|www-conf.slac.stanford.edu/snic/presentations/apr3/S04/snic_baron.pdf|


%\bibitem{reme85011}  F.Barbosa, V.Baturin, 
%V.Burkert, L.Elouadrhiri, C.Gueva, W.Kim, S.Majewski,
%G.Mutchler, V.Popov,  E.Smith,  C.Zorn.
% `` Time-of-flight start counter  with Burle 85011
% micro-channel plate photo-multipliers'', 
%to be published as CLAS NOTE, to be submitted to NIM.



\bibitem{r4} S. Taylor et al., ``The CLAS Start Counter'', Nucl. Instr. and Meth. A \textbf{462},484-493~(2001).


\end{thebibliography}
