%%
\input{../preshower/pcal-overview.tex}
\input{../preshower/design-analysis.tex}
%%\input{../preshower/expected-performance.tex}
\input{../preshower/mechdesign.tex}
\input{../preshower/pcal_readout.tex}
\input{../preshower/pcal_prototyp.tex}
\input{../preshower/pcal_construction.tex}
\input{../preshower/pcal_collab.tex}
%%
\begin{thebibliography}{99}

\bibitem{ec} M. Amarian {\it et al.}, NIM {\bf A 460}, 239 (2001),

\bibitem{natalia} N. Dashyan and S. Stepanyan, CLAS-Note 2007-001, 
%	{\tt http://www1.jlab.org/ul/Physics/Hall-B/clas/ }

\bibitem{whitlow} K. Whitlow, N. Dashyan and S. Stepanyan, 
	CLAS-Note 2007-002. 

\bibitem{tests} G. Asryan, H. Voskanyan, and S. Stepanyan, 
	CLAS-Note 2007-007. 

\bibitem{ecrec} N. Dashyan and S. Stepanyan, CLAS Note 2006-016.

\bibitem{philippe} Ph. Rosier, Preliminary Finite Element Analysis 
of the Preshower Calorimeter for the CLAS12 .

\bibitem{GROOVED} S. Stepanyan {\it et al.}, CLAS-NOTE 2002-001 (2002).

\bibitem{MINOS} MINOS Far detector, arXiv:hep-ex/0507018 (2005).

\bibitem{fnalex} FNAL Extruder http://www.nicadd.niu.edu/research/extruder.

\bibitem{yerex} M. Amarian {\it et al.}, CLAS-NOTE 1992-012 (1992).

\bibitem{yield} V. Burkert {\it et al.}, CLAS-NOTE 1992-008 (1992).

\bibitem{adapter} K. Giovanetti and A. Stavola, Fiber adapters for
pre-shower tests (JMU).

\bibitem{hbook} HBOOK, Statistical Analysis and Histogramming, CERN
program library long write-ups Y250.  

\bibitem{paw} Physics Analysis Workstation, CERN program library long
write-ups Q121.  

\end{thebibliography}
